---
title: "Data_manip"
author: "Mundi"
date: "2025-01-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Required packaged
```{r}
library(ggplot2)     # For data visualization
library(dplyr)       # For data manipulation (e.g., mutate, rename)
library(tidyr)       # For data tidying (e.g., handling missing values)
library(skimr)       # For quick data summaries
```

## Loading data and skimming

```{r}
load("C:/Users/Milan/Downloads/allgreI.RData")
load("C:/Users/Milan/Downloads/allgreD.RData")
load("C:/Users/Milan/Downloads/allgreM.RData")

skim(allgreI)
skim(allgreD)
skim(allgreM)
```

# Lets look at the structure of the data
```{r}
str(allgreI)
str(allgreD)
str(allgreM)
```
# Lets add our new variable, immobil to the dataset
```{r}
# Create a new variable 'immobil': 1 if `nbd` = 0 (immobile), otherwise 0
allgreI <- allgreI %>%
  mutate(immobil = ifelse(nbd == 0, 1, 0))

# Check if the variable was created correctly
table(allgreI$immobil)
```

# We will not be using all of these variables so lets only keep the ones we are interested in

```{r}
# Filter the dataset to include only the specified columns
allgreI_filtered <- allgreI %>% select(id_pers, immobil, cspgroup, sexe, dispovp, age)

# Verify the new filtered dataset
skim(allgreI_filtered)
```

# We see we have some missing values for dispovp which need to be handled. Since dispovp is never 0 we assume the NA values are to be set as 0 to indicate no available cars in the household.

```{r}
# Replace NA values in dispovp with 0
allgreI_filtered$dispovp[is.na(allgreI$dispovp)] <- 0

# Confirm that there are no NA values in VP_DISPO
sum(is.na(allgreI_filtered$dispovp))
```


# We want change the structure of age to group it instead of using it as a linear integer variable
```{r}
# Create age groups and remove the 'age' variable
allgreI_filtered <- allgreI_filtered %>%
  mutate(
    age_group = cut(
      age, 
      breaks = c(0, 20, 60, Inf),  # Breakpoints for age groups
      labels = c("0-20", "20-60", "60+"),
      right = FALSE  # Exclude the upper limit
    )
  ) %>%
  select(-age)  # Remove the 'age' column

# Verify that 'age' has been removed
str(allgreI_filtered)

```
```{r}
# Transformations and variable creations
allgreI_filtered <- allgreI_filtered %>%
  mutate(
    # Convert Y to a binary factor with meaningful labels
    immobil = factor(immobil, levels = c(0, 1), labels = c("No", "Yes")),
    
    # Convert cspgroup from integer to factor
    cspgroup = factor(cspgroup),
    
    # Add a binary variable indicating whether there is at least one car
    has_car = ifelse(dispovp > 0, 1, 0),
    
    # Convert has_car to a factor for binary interpretation
    has_car = factor(has_car, levels = c(0, 1), labels = c("No", "Yes")),
    
    # Convert sexe to a factor with labels Male and Female
    sexe = factor(sexe, levels = c(1, 2), labels = c("Male", "Female"))
  )

# Verify the changes
str(allgreI_filtered)
```
# For cspgroup we group the irrelevant groups (based on logit model 3)
```{r}
allgreI_filtered <- allgreI_filtered %>%
  mutate(
    cspgroup_combined = case_when(
      cspgroup %in% c("2", "5", "6", "8", "9") ~ "non_significant", # Niet-significante groepen
      TRUE ~ as.character(cspgroup) # Overige groepen behouden
    )
  )

# Zet de nieuwe variabele om in een factor
allgreI_filtered$cspgroup_combined <- as.factor(allgreI_filtered$cspgroup_combined)

# Controleer de verdeling van de nieuwe variabele
cat("Distribution of cspgroup_combined:\n")
print(table(allgreI_filtered$cspgroup_combined))

# Verwijder 'cspgroup' en voeg 'cspgroup_combined' toe aan allgreI
allgreI_filtered <- allgreI_filtered %>%
  select(-cspgroup) %>%    # Verwijder de kolom 'cspgroup'
  rename(cspgroup = cspgroup_combined)  # Hernoem 'cspgroup_combined' naar 'cspgroup'

# Controleer of de wijziging correct is doorgevoerd
str(allgreI_filtered)


```


# Now we save the manipulated data into a new RData file to work with in the modeling
```{r}
# Save the transformed dataset to an RData file
save(allgreI_filtered, file = "allgreI_filtered.RData")
```

# Now we create a W variable
# We start by joining the datasets together
```{r}
# Datasets samenvoegen op basis van 'id_pers'
merged_data <- allgreM %>%
  inner_join(allgreI, by = c("id_men", "id_pers")) %>%
  inner_join(allgreD, by = c("id_men", "id_pers", "id_depl"))


# Controleer de eerste paar rijen van de samengevoegde dataset
head(merged_data)
```
# Now we look at the missing values of the relevant variables to create our W variable
```{r}
# Analyze missing values in the relevant variables
missing_summary <- merged_data %>%
  select(NB_velo, dispovp, ABO_TC, abonpeage, LIEU_STAT, NB_2Rm) %>%
  summarise_all(~sum(is.na(.)))

# Print the summary of missing values
print(missing_summary)

# Visualize the distribution of each variable
merged_data_long <- merged_data %>%
  select(NB_velo, dispovp, ABO_TC, abonpeage, LIEU_STAT) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value")

# Plot the distribution per variable
ggplot(merged_data_long, aes(x = as.factor(Value))) +
  geom_bar() +
  facet_wrap(~Variable, scales = "free") +
  labs(title = "Distribution of Variables",
       x = "Value",
       y = "Frequency") +
  theme_minimal()

# Alternative visualization of missing values with 'naniar' (optional)
# library(naniar)
# vis_miss(merged_data %>% select(NB_velo, VP_DISPO, dispovp, ABO_TC, abonpeage, LIEU_STAT))

# New chunk: Convert NA to 0 for 'abonpeage', 'dispovp', and 'LIEU_STAT'
merged_data <- merged_data %>%
  mutate(
    abonpeage = replace_na(abonpeage, 0),
    dispovp = replace_na(dispovp, 0),
    LIEU_STAT = replace_na(LIEU_STAT, 0)
  )

# Check if the conversion was successful
missing_summary_updated <- merged_data %>%
  select(abonpeage, dispovp, LIEU_STAT) %>%
  summarise_all(~sum(is.na(.)))

print(missing_summary_updated)

```
# Now we create the W variable
```{r}
# New chunk: Create variable 'W'
merged_data <- merged_data %>%
  mutate(W = ifelse(NB_velo == 0 & NB_2Rm == 0 & dispovp == 0 & ABO_TC == 3 & abonpeage == 0 & LIEU_STAT == 0, 0, 1))

# Check the distribution of 'W' with detailed output
cat("Distribution of W:\n")
print(table(merged_data$W))

# Plot the distribution of 'W'
ggplot(merged_data, aes(x = as.factor(W))) +
  geom_bar(fill = "steelblue") +
  labs(title = "Distribution of W",
       x = "W Value",
       y = "Frequency") +
  theme_minimal()

```
```{r}
# Merge 'W' into allgreI_filtered using 'id_pers'
allgreI_filtered <- allgreI_filtered %>%
  left_join(merged_data %>% select(id_pers, W), by = "id_pers")

# Verify that 'W' was correctly added
str(allgreI_filtered)

# Save the updated dataset with 'W' included
save(allgreI_filtered, file = "allgreI_with_W.RData")

# Optional: View the first few rows to check
head(allgreI_filtered)

```
# Check if W is joint well
```{r}
# Check for any rows where dispovp > 0 but W is not 1
invalid_cases <- allgreI_filtered %>%
  filter(dispovp > 0 & W != 1)

# Display the number of violations
cat("Number of cases where dispovp > 0 and W is not 1:", nrow(invalid_cases), "\n")

# If any invalid cases exist, display them
if(nrow(invalid_cases) > 0) {
  print(invalid_cases)
} else {
  cat("Validation passed: W is always 1 when dispovp > 0.\n")
}

```

```{r}
selected_vars <- c("TYPE_HAB", "zoneres.x", "PBM_STAT", "STAT_TRAV", "OCCU1", "travdom", "nb_pers")

# Selectie van de variabelen uit de gejoinede dataset
filtered_data <- merged_data %>% select(all_of(selected_vars))

# Bekijk de algemene samenvatting van de geselecteerde variabelen
skim(filtered_data)

# Distributie van elke variabele + aantal ontbrekende waarden visualiseren
for (var in selected_vars) {
  print(ggplot(filtered_data, aes_string(x = var)) +
          geom_bar(na.rm = TRUE, fill = "steelblue") +
          ggtitle(paste("Distributie van", var)) +
          theme_minimal())
  
  # Aantal missende waarden tonen
  missing_count <- sum(is.na(filtered_data[[var]]))
  total_count <- nrow(filtered_data)
  cat(paste("Variabele:", var, "\nOntbrekende waarden:", missing_count, 
            "\nPercentage ontbrekend:", round((missing_count / total_count) * 100, 2), "%\n\n"))
}
```

```{r}
# Combineer alle mutaties in één blok
merged_data <- merged_data %>%
  mutate(
    # Vervang NA door 0
    PBM_STAT = replace_na(PBM_STAT, 0),
    STAT_TRAV = replace_na(STAT_TRAV, 0),
    
    # Zet de geselecteerde variabelen om naar factor (categorisch)
    TYPE_HAB = as.factor(TYPE_HAB),
    zoneres.x = as.factor(zoneres.x),
    OCCU1 = as.factor(OCCU1),
    travdom = as.factor(travdom),
    nb_pers = as.factor(nb_pers),
    
    # Creëer de binaire variabele 'parking_diff'
    parking_diff = ifelse(PBM_STAT == 1 | STAT_TRAV == 1, 1, 0)
  ) %>%
  select(-PBM_STAT, -STAT_TRAV)  # Verwijder PBM_STAT en STAT_TRAV

# Voeg de aangepaste variabelen toe aan allgreI_filtered via left_join
allgreI_filtered <- allgreI_filtered %>%
  left_join(merged_data %>% select(id_pers, TYPE_HAB, zoneres.x, OCCU1, travdom, nb_pers, parking_diff),
            by = "id_pers")

# Controleer of de variabelen correct zijn toegevoegd
str(allgreI_filtered)

# Opslaan van de geüpdatete dataset
save(allgreI_filtered, file = "allgreI_with_new_vars.RData")

# Bekijk een overzicht van de eerste paar rijen
head(allgreI_filtered)

```
```{r}
library(dplyr)

# 1. Vind alle kolommen die eindigen op .x of .y
suffix_columns <- names(allgreI_filtered)[grepl("\\.x$|\\.y$", names(allgreI_filtered))]

# 2. Loop door alle suffix kolommen en verwerk ze
for (col in suffix_columns) {
  # Haal de originele kolomnaam zonder suffix
  base_col <- gsub("\\.x$|\\.y$", "", col)
  
  # Controleer of de kolom zonder suffix al bestaat
  if (base_col %in% names(allgreI_filtered)) {
    # Vergelijk de kolommen
    if (identical(allgreI_filtered[[col]], allgreI_filtered[[base_col]])) {
      # Als ze identiek zijn, verwijder de .x/.y kolom
      allgreI_filtered <- allgreI_filtered %>%
        select(-all_of(col))
      cat("✅", col, "is identiek aan", base_col, "en is verwijderd.\n")
    } else {
      cat("⚠️", col, "verschilt van", base_col, ". Handmatige controle vereist.\n")
    }
  } else {
    # Als de basisversie niet bestaat, hernoem de kolom zonder suffix
    allgreI_filtered <- allgreI_filtered %>%
      rename(!!base_col := all_of(col))
    cat("🔄", col, "is hernoemd naar", base_col, ".\n")
  }
}

# 3. Verwijder PBM_STAT en STAT_TRAV als ze nog bestaan
allgreI_filtered <- allgreI_filtered %>%
  select(-any_of(c("PBM_STAT", "STAT_TRAV")))

cat("🗑️  PBM_STAT en STAT_TRAV zijn verwijderd.\n")

# 4. Controleer of er nog kolommen met .x of .y zijn
remaining_suffixes <- names(allgreI_filtered)[grepl("\\.x$|\\.y$", names(allgreI_filtered))]

if (length(remaining_suffixes) == 0) {
  cat("🎉 Alle kolommen met .x en .y zijn succesvol verwijderd of hernoemd!\n")
} else {
  cat("❗ Er zijn nog kolommen met .x of .y:\n")
  print(remaining_suffixes)
}

# 5. Controleer de structuur van de opgeschoonde dataset
str(allgreI_filtered)

```


```{r}
# Zet W en parking_diff om naar binaire factorvariabelen met labels "No" en "Yes"
allgreI_filtered <- allgreI_filtered %>%
  mutate(
    W = factor(W, levels = c(0, 1), labels = c("No", "Yes")),
    parking_diff = factor(parking_diff, levels = c(0, 1), labels = c("No", "Yes"))
  )

# Controleer of de omzetting correct is uitgevoerd
```


```{r}
str(allgreI_filtered)
```
```{r}
library(ggplot2)
library(dplyr)

# 1. Splits categorische en numerieke variabelen
cat_vars <- names(allgreI_filtered)[sapply(allgreI_filtered, is.factor)]
num_vars <- names(allgreI_filtered)[sapply(allgreI_filtered, is.numeric)]

# 2. Distributies van categorische variabelen
cat("📊 Distributie van categorische variabelen:\n")
for (var in cat_vars) {
  cat("🔸", var, "\n")
  
  # Frequentietabel printen
  print(table(allgreI_filtered[[var]], useNA = "ifany"))
  
  # Staafdiagram maken
  plot <- ggplot(allgreI_filtered, aes_string(x = var)) +
    geom_bar(fill = "steelblue", na.rm = TRUE) +
    labs(title = paste("Distributie van", var), x = var, y = "Frequentie") +
    theme_minimal()
  
  print(plot)
}

# 3. Distributies van numerieke variabelen
cat("\n📈 Distributie van numerieke variabelen:\n")
for (var in num_vars) {
  cat("🔹", var, "\n")
  
  # Samenvatting printen
  print(summary(allgreI_filtered[[var]]))
  
  # Histogram maken
  plot <- ggplot(allgreI_filtered, aes_string(x = var)) +
    geom_histogram(binwidth = 1, fill = "steelblue", color = "black", na.rm = TRUE) +
    labs(title = paste("Distributie van", var), x = var, y = "Frequentie") +
    theme_minimal()
  
  print(plot)
}

```
```{r}
library(dplyr)
library(ggplot2)

# TYPE_HAB correct groeperen met grondige schoonmaak
allgreI_filtered <- allgreI_filtered %>%
  mutate(
    TYPE_HAB = case_when(
      TYPE_HAB %in% c("1", "2") ~ "Groep 1",      # Groep 1: 1 of 2
      TYPE_HAB %in% c("3", "4", "5") ~ "Groep 2",   # Groep 2: 3, 4, 5                 # Overige waarden worden NA
    ),
    TYPE_HAB = factor(TYPE_HAB, levels = c("Groep 1", "Groep 2"))
  )


# 2️⃣ OCCU1 correct groeperen
allgreI_filtered <- allgreI_filtered %>%
  mutate(
    OCCU1 = case_when(
      OCCU1 %in% c("3", "4", "5") ~ "3-5",  # Waarden 3, 4, 5 samengevoegd
      TRUE ~ as.character(OCCU1)
    ),
    OCCU1 = factor(OCCU1)
  )

# 3️⃣ Nieuwe variabele 'retrait' toevoegen
allgreI_filtered <- allgreI_filtered %>%
  mutate(
    retrait = ifelse(OCCU1 == "7", "Yes", "No"),
    retrait = factor(retrait, levels = c("No", "Yes"))
  )

# 2️⃣ nb_pers correct groeperen
allgreI_filtered <- allgreI_filtered %>%
  mutate(
    nb_pers = as.numeric(as.character(nb_pers)),  # Zet om naar numeriek
    nb_pers = case_when(
      nb_pers %in% c(1, 2) ~ as.character(nb_pers),
      nb_pers %in% c(3, 4) ~ "3-4",  # Groep 3-4
      nb_pers %in% c(5, 6, 7, 8) ~ "5-8",  # Groep 5-8
      TRUE ~ NA_character_
    ),
    nb_pers = factor(nb_pers, levels = c("1", "2", "3-4", "5-8"))
  )

# 🔍 Controleer de aangepaste dataset
str(allgreI_filtered)

# ℹ️ Bekijk de verdeling van de nieuwe groepen
cat("Verdeling van TYPE_HAB:\n")
print(table(allgreI_filtered$TYPE_HAB, useNA = "ifany"))

cat("Verdeling van OCCU1:\n")
print(table(allgreI_filtered$OCCU1, useNA = "ifany"))

cat("Verdeling van retrait:\n")
print(table(allgreI_filtered$retrait, useNA = "ifany"))

cat("Verdeling van nb_pers:\n")
print(table(allgreI_filtered$nb_pers, useNA = "ifany"))

# 📊 5️⃣ Visualisatie van de distributies van alle variabelen
cat_vars <- names(allgreI_filtered)[sapply(allgreI_filtered, is.factor)]
num_vars <- names(allgreI_filtered)[sapply(allgreI_filtered, is.numeric)]

# 📊 Visualisatie voor categorische variabelen
for (var in cat_vars) {
  cat("🔸 Distributie van", var, ":\n")
  
  # Frequentietabel printen
  print(table(allgreI_filtered[[var]], useNA = "ifany"))
  
  # Staafdiagram maken
  plot <- ggplot(allgreI_filtered, aes_string(x = var)) +
    geom_bar(fill = "steelblue", na.rm = TRUE) +
    labs(title = paste("Distributie van", var), x = var, y = "Frequentie") +
    theme_minimal()
  
  print(plot)
}

# 📈 Visualisatie voor numerieke variabelen
for (var in num_vars) {
  cat("🔹 Distributie van", var, ":\n")
  
  # Samenvatting printen
  print(summary(allgreI_filtered[[var]]))
  
  # Histogram maken
  plot <- ggplot(allgreI_filtered, aes_string(x = var)) +
    geom_histogram(binwidth = 1, fill = "steelblue", color = "black", na.rm = TRUE) +
    labs(title = paste("Distributie van", var), x = var, y = "Frequentie") +
    theme_minimal()
  
  print(plot)
}

```
