---
title: "Untitled"
author: "Mundi"
date: "2024-10-10"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(reshape2)
library(corrplot)
library(skimr)
library(tidyr)
```

## Slide with Plot

```{r}
load("C:/Users/Milan/Downloads/allgreD.RData")
load("C:/Users/Milan/Downloads/allgreI.RData")
load("C:/Users/Milan/Downloads/allgreM.RData")
```

```{r}
skim(allgreD)
```

```{r}
skim(allgreI)

```
```{r}
skim(allgreM)
```


```{r}
summary(allgreI)


```

```{r}
summary(allgreM)
```
```{r}
# Function to return the class or type of each column
column_types <- function(allgreD) {
  sapply(allgreD, function(column) class(column)[1])
}

# Get the type for each column
column_types(allgreD)
```
```{r}
head(allgreI$nbd)
sum(is.na(allgreI$nbd))

ggplot(allgreI, aes(x = nbd)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Verdeling van de variabele 'nbd'", x = "nbd", y = "Frequentie")
```
```{r}
# Zorg ervoor dat dplyr geladen is
library(dplyr)

# Creëer de nieuwe variabele 'immobil' op basis van de waarde van 'nbd'
allgreI <- allgreI %>%
  mutate(immobil = ifelse(nbd == 0, 1, 0))

# Controleer de eerste paar rijen om te zien of de variabele goed is aangemaakt
allgreI[, c("nbd", "immobil")]

# Tel het aantal rijen waar immobil gelijk is aan 1
sum(allgreI$immobil == 1)


```

```{r}
allgreI[, c("X", "Y", "Z")]  # Vervang door de juiste kolomnamen als nodig

```
```{r}
allgreI_clean <- na.omit(allgreI)
allgreI$Z[is.na(allgreI$Z)] <- mean(allgreI$Z, na.rm = TRUE)
allgreI$Y <- as.factor(allgreI$Y)  # Voor logistische regressie
allgreI$X <- as.numeric(allgreI$X)
allgreI$Z <- as.numeric(allgreI$Z)
```

```{r}
# Logistische regressie: Y is binaire uitkomst, X en Z zijn predictorvariabelen
model <- glm(Y ~ X + Z, data = allgreI, family = binomial)

# Bekijk de samenvatting van het model
summary(model)

```
## Conclusion: The younger people are, the more immobile they are, number of cars is irrelevant, only a small part of the variability is declared in this model

