---
title: "Mobility behaviour"
output: html_document
date: "2025-01-06"
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
# Required Packages
library(ggplot2)     # For data visualization
library(dplyr)       # For data manipulation (e.g., mutate, rename)
library(tidyr)       # For data tidying (e.g., handling missing values)
library(skimr)       # For quick data summaries

```

## Loading and exploring


```{r cars}
# Load the necessary dataset
load("C:/Users/Milan/Downloads/allgreI.RData")

# Check the structure of the dataset
str(allgreI)

# Summary of the dataset
summary(allgreI)
head(allgreI)
skim(allgreI)

# Check for missing values
sapply(allgreI, function(x) sum(is.na(x)))

# View the distribution of the variable `nbd` (number of trips)
table(allgreI$nbd)

# Visualize the distribution of `nbd`
library(ggplot2)
ggplot(allgreI, aes(x = nbd)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Distribution of Number of Trips (nbd)", x = "Number of Trips", y = "Frequency")

```
# If we want to clean all of the data, optional

```{r}
# Identify missing values per variable, also percentage wise


# Count missing values for each variable
missing_values <- sapply(allgreI_clean, function(x) sum(is.na(x)))

# Display variables with missing values
missing_values[missing_values > 0]

# Alternatively, show percentage of missing values
missing_percent <- sapply(allgreI_clean, function(x) mean(is.na(x)) * 100)
missing_percent[missing_percent > 0]

# Identify variables with less than 10% missing values
missing_percent <- sapply(allgreI_clean, function(x) mean(is.na(x)) * 100)
vars_less_10 <- names(missing_percent[missing_percent > 0 & missing_percent <= 10])

# Calculate the number of rows remaining after removing rows with missing values in these variables
rows_remaining <- nrow(allgreI_clean[complete.cases(allgreI_clean[vars_less_10]), ])

# Print the result
cat("Number of rows remaining after removing rows with missing values in variables with <10% missing values:", 
    rows_remaining, "out of", nrow(allgreI_clean), "\n")

# we can delete these rows
# these variables are: TEL_PORT, mail, permis, etabscol, OCCU1, OCCU2, csp, ABO_TC, fqvelo, FQ2R1, FQ2R2, fqvpcond, fqvppass, freqtcu, freqtram, freqrurb, freqtransisere, freqter, statut2

# Remove rows with missing values in variables with <10% missing values
allgreI_clean <- allgreI_clean[complete.cases(allgreI_clean[vars_less_10]), ]

# Confirm remaining rows
cat("Number of rows after removing rows with missing values:", nrow(allgreI_clean), "\n")

# 10% - 50% -> we start by using the medium, but later we will check whether they are relevant, to see what the best way of handling these is
# these variable are: btt (16.3%), travdom (35.2%), zonetrav (37.6%)

# For numerical variable `btt`: Impute missing values with the median
allgreI_clean$btt[is.na(allgreI_clean$btt)] <- median(allgreI_clean$btt, na.rm = TRUE)

# Helper function to calculate the mode for categorical variables
mode_impute <- function(x) {
  uniq_vals <- na.omit(x)
  return(uniq_vals[which.max(tabulate(match(uniq_vals, uniq_vals)))])
}

# For categorical variable `travdom`: Impute missing values with the mode
allgreI_clean$travdom[is.na(allgreI_clean$travdom)] <- mode_impute(allgreI_clean$travdom)

# For categorical variable `zonetrav`: Impute missing values with the mode
allgreI_clean$zonetrav[is.na(allgreI_clean$zonetrav)] <- mode_impute(allgreI_clean$zonetrav)

# Optional: Verify that there are no more missing values in these variables
sapply(allgreI_clean[, c("btt", "travdom", "zonetrav")], function(x) sum(is.na(x)))

# >50% -> exclude
# these variable are: VAL_ABO (78%), dispovp (56.4%), PBM_STAT (68.3%), STAT_TRAV (88.1%), situveil (90.9%)

# Drop columns with more than 50% missing data
missing_percent <- sapply(allgreI_clean, function(x) mean(is.na(x)) * 100)
allgreI_clean <- allgreI_clean[, missing_percent <= 50]

# Count the total missing values for each variable
missing_values <- sapply(allgreI_clean, function(x) sum(is.na(x)))

# Display variables with missing values (should return an empty result if all missing values are treated)
missing_values[missing_values > 0]

# Alternatively, show the percentage of missing values
missing_percent <- sapply(allgreI_clean, function(x) mean(is.na(x)) * 100)

# Display variables with missing percentages greater than 0
missing_percent[missing_percent > 0]

summary(allgreI_clean)

# Check distributions of key variables
table(allgreI_clean$nbd)        # Distribution of number of trips
table(allgreI_clean$immobil)    # Consistency of immobility variable

cat("Number of rows:", nrow(allgreI_clean), "\n")
cat("Number of columns:", ncol(allgreI_clean), "\n")

```

## Creating the immobil variable

```{r pressure, echo=FALSE}
# Create a new variable 'immobil': 1 if `nbd` = 0 (immobile), otherwise 0
allgreI <- allgreI %>%
  mutate(immobil = ifelse(nbd == 0, 1, 0))

# Check if the variable was created correctly
table(allgreI$immobil)

```

```{r}
# Distribution of 'immobil'
table(allgreI$immobil)

# Percentage distribution
prop.table(table(allgreI$immobil)) * 100

# Visualization of 'immobil'
library(ggplot2)
ggplot(allgreI, aes(x = factor(immobil))) +
  geom_bar(fill = "steelblue") +
  labs(title = "Distribution of Mobility Behavior", x = "Immobile (1 = Yes, 0 = No)", y = "Count")

```
# Now let's look at age
```{r}
# Summary statistics
summary(allgreI$age)

# Histogram of age
ggplot(allgreI, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  labs(title = "Distribution of Age", x = "Age", y = "Frequency")

```
# Relationship between age and immobil
```{r}
# Boxplot of age by immobil
ggplot(allgreI, aes(x = factor(immobil), y = age)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Age vs Mobility Behavior", x = "Immobile (1 = Yes, 0 = No)", y = "Age")

```
# Now we look at VP_dispo
```{r}
colnames(allgreI)[colnames(allgreI) == "dispovp"] <- "VP_DISPO"

# Replace NA values in VP_DISPO with 0
allgreI$VP_DISPO[is.na(allgreI$VP_DISPO)] <- 0

# Confirm that there are no NA values in VP_DISPO
sum(is.na(allgreI$VP_DISPO))


# Summary statistics
table(allgreI$VP_DISPO)

# Barplot of car ownership
ggplot(allgreI, aes(x = factor(VP_DISPO))) +
  geom_bar(fill = "orange") +
  labs(title = "Car Ownership Distribution", x = "Number of Cars", y = "Count")

```

# Relationship between available cars and immobil
```{r}
# Barplot of car ownership by mobility behavior
ggplot(allgreI, aes(x = factor(VP_DISPO), fill = factor(immobil))) +
  geom_bar(position = "dodge") +
  labs(title = "Car Ownership vs Mobility Behavior", x = "Number of Cars", y = "Count", fill = "Immobile")

```

```{r}
# Rename variables
allgreI <- allgreI %>%
  rename(Y = immobil, X = age, Z = VP_DISPO)
```
```{r}
# Check for NA values in the renamed variables
na_check <- sapply(allgreI[, c("X", "Y", "Z")], function(x) sum(is.na(x)))
print(na_check)

```

# Now we can start doing linear regressions
```{r}
# First regression: Y ~ X
model1 <- lm(Y ~ X, data = allgreI)
summary(model1)

# Second regression: Y ~ Z
model2 <- lm(Y ~ Z, data = allgreI)
summary(model2)

# Third regression: Y ~ X + Z
model3 <- lm(Y ~ X + Z, data = allgreI)
summary(model3)

# Fourth regression: Y ~ X * Z (including interaction)
model4 <- lm(Y ~ X * Z, data = allgreI)
summary(model4)

```
# We can see that both variables are significant. Age has a very small negative coëfficiënt. The number of cars has a bigger negative coëfficiënt and declares more of the variability. Their interaction is slightly significant and has a very small negative coëfficiënt.

# Since immobil is a categorical variable, a logistic regression will ofcourse be better suited.

```{r}
# Controleer het datatype van X en Z
class(allgreI$X)
class(allgreI$Z)

# Alternatief: Controleer of X en Z als factor zijn gedefinieerd
is.factor(allgreI$X)
is.factor(allgreI$Z)

# Zet X en Z om naar factoren
allgreI$X <- as.factor(allgreI$X)
allgreI$Z <- as.factor(allgreI$Z)

# Controleer opnieuw het datatype
class(allgreI$X)
class(allgreI$Z)

```


```{r}

# First logistic regression: Y ~ X
logit_model1 <- glm(Y ~ X, data = allgreI, family = binomial)
summary(logit_model1)

# Second logistic regression: Y ~ Z
logit_model2 <- glm(Y ~ Z, data = allgreI, family = binomial)
summary(logit_model2)

```
# We get similar coëfficients, but now we can see that the AIC for the model with number of cars is lower than the model with only age. The model with both performs slightly better, but this model is less simple. The model with the interaction performs the best of the 4, but is also the most complicated one.

```{r}
# Controleer de structuur van X
str(allgreI$X)

# Als X een factor is, zet deze om naar een numerieke variabele
allgreI$X_numeric <- as.numeric(as.character(allgreI$X))

# Controleer de nieuwe numerieke variabele
summary(allgreI$X_numeric)

# Maak leeftijdsgroepen aan met de numerieke versie van X
allgreI <- allgreI %>%
  mutate(
    age_group = cut(
      X_numeric, 
      breaks = c(0, 16, 20, 60, 75, 85, Inf), # Breaks voor leeftijdsgroepen
      labels = c("0-16", "16-20", "20-60", "60-75", "75-85", "85+"),
      right = FALSE # Sluit de bovenste grens uit
    )
  )

# Controleer of de nieuwe variabele correct is aangemaakt
table(allgreI$age_group)

# Zet de nieuwe variabele om naar een factor voor regressie
allgreI$age_group <- as.factor(allgreI$age_group)

# Verwijder de tijdelijke numerieke variabele als je deze niet meer nodig hebt
allgreI$X_numeric <- NULL


```
```{r}
logit_model1 <- glm(Y ~ age_group, data = allgreI, family = binomial)
summary(logit_model1)
```
```{r}
# Origineel model
model_full <- glm(Y ~ age_group, family = binomial, data = allgreI)

# Nieuw model met samengevoegde groepen
allgreI <- allgreI %>%
  mutate(
    age_group_combined = case_when(
      age_group %in% c("25-35", "35-45", "45-55") ~ "25-55",
      TRUE ~ as.character(age_group)
    )
  )

# Zet nieuwe variabele om naar een factor
allgreI$age_group_combined <- as.factor(allgreI$age_group_combined)

# Nieuw model
model_reduced <- glm(Y ~ age_group_combined, family = binomial, data = allgreI)

# Likelihood Ratio Test
anova(model_full, model_reduced, test = "LRT")

```
```{r}
logit_model1 <- glm(Y ~ age_group+Z, data = allgreI, family = binomial)
summary(logit_model1)
logit_model1 <- glm(Y ~ age_group*Z, data = allgreI, family = binomial)
summary(logit_model1)
```

```{r}
# Remove rows where ABO_TC is NA
allgreI <- allgreI[!is.na(allgreI$ABO_TC), ]

# Convert ABO_TC to a factor
allgreI$ABO_TC <- factor(allgreI$ABO_TC, levels = c(1, 2, 3), 
                         labels = c("Free", "Paid", "None"))

# Check if NAs have been successfully removed
cat("Unique values in ABO_TC after removing NAs:\n")
print(levels(allgreI$ABO_TC))

# Show the distribution of ABO_TC without NAs
cat("\nDistribution of ABO_TC without NAs:\n")
print(table(allgreI$ABO_TC))

# Visualize the distribution of ABO_TC without NAs
ggplot(allgreI, aes(x = ABO_TC)) +
  geom_bar(fill = "steelblue", color = "black") +
  labs(
    title = "Distribution of ABO_TC (without NAs)",
    x = "Type of subscription",
    y = "Count"
  ) +
  theme_minimal()


```

```{r}
logit_model1 <- glm(Y ~ ABO_TC, data = allgreI, family = binomial)
summary(logit_model1)
logit_model1 <- glm(Y ~ age_group + ABO_TC + Z, data = allgreI, family = binomial)
summary(logit_model1)
```

```{r}
# Check if cspgroup is already a factor
if (!is.factor(allgreI$cspgroup)) {
  # Convert cspgroup to a factor
  allgreI$cspgroup <- as.factor(allgreI$cspgroup)
}

# Check the unique levels in cspgroup
cat("Levels in cspgroup:\n")
print(levels(allgreI$cspgroup))

# Check the distribution of cspgroup
cat("\nDistribution of cspgroup:\n")
print(table(allgreI$cspgroup))

# Check for missing values in csp
cat("Number of missing values in csp:\n")
print(sum(is.na(allgreI$cspgroup)))

# Convert csp to a factor (if it is not already)
if (!is.factor(allgreI$cspgroup)) {
  allgreI$csp <- as.factor(allgreI$cspgroup)
}

# Show the distribution of csp
cat("\nDistribution of csp:\n")
print(table(allgreI$cspgroup))

# Visualize the distribution of csp
library(ggplot2)
ggplot(allgreI, aes(x = csp)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(
    title = "Distribution of CSPgroup",
    x = "CSP Categories",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for better readability

```

```{r}
logit_model1 <- glm(Y ~ cspgroup, data = allgreI, family = binomial)
summary(logit_model1)
logit_model1 <- glm(Y ~ cspgroup + age_group + Z, data = allgreI, family = binomial)
summary(logit_model1)
```
```{r}
# Maak een nieuwe variabele waarin niet-significante groepen worden samengevoegd
allgreI <- allgreI %>%
  mutate(
    cspgroup_combined = case_when(
      cspgroup %in% c("2", "5", "6", "8", "9") ~ "non_significant", # Niet-significante groepen
      TRUE ~ as.character(cspgroup) # Overige groepen behouden
    )
  )

# Zet de nieuwe variabele om in een factor
allgreI$cspgroup_combined <- as.factor(allgreI$cspgroup_combined)

# Controleer de verdeling van de nieuwe variabele
cat("Distribution of cspgroup_combined:\n")
print(table(allgreI$cspgroup_combined))

# Voer een nieuw model uit met de samengevoegde groepen
logit_model <- glm(Y ~ cspgroup_combined, family = binomial, data = allgreI)
summary(logit_model)

logit_model1 <- glm(Y ~ cspgroup_combined + age_group + Z, data = allgreI, family = binomial)
summary(logit_model1)
```

