---
title: "Mobility behavior of mobile and immobile people"
author: "G. Arnarson, F. Bartke, M. van Leare, S. Nguyen"
date: "`r Sys.Date()`"
output: beamer_presentation
theme: "Antibes"
colortheme: "beaver" 
header-includes:
- |
    \setbeamertemplate{footline}{
      \leavevmode%
      \hbox{\hspace*{0.3cm}G. Arnarson, F. Bartke, M. van Leare, S. Nguyen \hfill \hspace*{7cm}\insertframenumber{} / \inserttotalframenumber}%
      \vskip0pt%
    }
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Motivation/ Introduction



## Agenda
1. Motivation/ Introduction
\pause
2. Agenda
\pause
3. Problem definition of the project
\pause
4. Exploratory Data Analysis
\pause
5. Dataset preparation
\pause
6. Modeling Techniques
\pause
7. Results and Interpretation
\pause
8. Challenges and Limitations
\pause
9. Conclusion and Outlook
\pause
10. Literature used
\pause
11. 
\pause


## Problem definition of the project


## Exploratory Data Analysis 

-   Bullet 1
-   Bullet 2
-   Bullet 3


## Dataset preparation 


## Modeling Techniques
Models used for analysis:
1. Logistic Regression
2. Decision Trees
3. Random Forest
4. Neural Networks 

```{r cars, echo = TRUE}
summary(cars)
```

## Results and Interpretation

performance?

```{r pressure}
plot(pressure)
```

## Challenges and Limitations



## Conclusion and Outlook 



## Thank you for your attention

\LARGE Are there any questions?





