---
title: "Mobility behavior of mobile and immobile people"
author: "G. Arnarson, F. Bartke, M. van Leare, S. Nguyen"
date: "`r Sys.Date()`"
output: beamer_presentation
theme: "Antibes"
colortheme: "beaver" 
header-includes:
- |
    \setbeamertemplate{footline}{
      \leavevmode%
      \hbox{\hspace*{0.3cm}G. Arnarson, F. Bartke, M. van Leare, S. Nguyen \hfill \hspace*{7cm}\insertframenumber{} / \inserttotalframenumber}%
      \vskip0pt%
    }
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Motivation/ Introduction



## Agenda
1. Motivation/ Introduction
\pause
2. Agenda
\pause
3. Problem definition of the project
\pause
4. Exploratory Data Analysis
\pause
5. Dataset preparation
\pause
6. Modeling Techniques
\pause
7. Results and Interpretation
\pause
8. Challenges and Limitations
\pause
9. Conclusion and Outlook
\pause
10. Literature used
\pause
11. 
\pause


## Problem definition of the project
Mobility behavior
-   Characterisation
-   Variables discussion

## Literature review and relevant variables

-   Literature: age, working status, health, sex, household, income, possession of a car, living area and space, distance to shops
-   Common sense / what we thought: spc, remote or home work, possession of bike or motorcycle, distance to public transport
-   Variables we consider: id_pers, immobil, cspgroup, sexe, dispovp, age, NO_MEN, TYPE_HAB, NB_2Rm, zoneres.x, Gare2, PBM_STAT, STAT_TRAV


## Dataset preparation 
-   Merging the datasets
-   Selecting the relevant variables
-   Factorizing, redefining and grouping categories
-   Defining and adding the W variable

## Data exploration
-   Initial exploration: Chi-square test and t-test
-> All of them, except zoneres are significant
-   Looking at collinearity through vif (has_car and travdom has mild collinearity)

## Logit model
-   Investigate which variables to use for a model, needs more grouping
## Modeling Techniques
Models used for analysis:
1. Logistic Regression
2. Decision Trees
3. Random Forest
4. Neural Networks 

```{r cars, echo = TRUE}
summary(cars)
```

## Results and Interpretation

performance?

```{r pressure}
plot(pressure)
```

## Challenges and Limitations



## Conclusion and Outlook 



## Thank you for your attention

\LARGE Are there any questions?





